---
title: "Living cover crop- Github"
output: html_document
date: "2026-09-24"
---

```{r setup, include=FALSE}
# ============================================================
# Report settings
# ============================================================

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = TRUE,
  cache = FALSE,
  fig.align = "center",
  out.width = "100%",
  dpi = 160
)

# ============================================================
# Install and load packages
# ============================================================

packages <- c("readxl", "dplyr", "tidyr", "ggplot2")

for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(
      pkg,
      repos = "https://cloud.r-project.org"
    )
  }
}

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)

# ============================================================
# Download the workbook from the public GitHub repository
# ============================================================

github_url <- paste0(
  "https://raw.githubusercontent.com/",
  "Pr1nc3sso/Xilei-Living-cover-crop/",
  "master/2-Data/Raw/Raw/",
  "Broccoli%20harvest.xlsx"
)

file_path <- tempfile(
  pattern = "Broccoli_harvest_",
  fileext = ".xlsx"
)

options(timeout = max(120, getOption("timeout", 60)))

download_status <- tryCatch(
  utils::download.file(
    url = github_url,
    destfile = file_path,
    method = "libcurl",
    mode = "wb",
    quiet = TRUE
  ),
  error = function(e) {
    stop(
      paste(
        "Could not download the Excel workbook.",
        "Check the repository visibility, file path,",
        "branch name, and internet connection.",
        conditionMessage(e)
      ),
      call. = FALSE
    )
  }
)

if (download_status != 0 ||
    !file.exists(file_path) ||
    file.info(file_path)$size == 0) {
  stop(
    "The workbook download failed or returned an empty file.",
    call. = FALSE
  )
}

# Save results beside the R Markdown document.
output_dir <- file.path(
  getwd(),
  "Broccoli_R_results"
)

dir.create(
  output_dir,
  showWarnings = FALSE,
  recursive = TRUE
)

# ============================================================
# Read and clean the data
# ============================================================

raw <- readxl::read_excel(
  path = file_path,
  sheet = "living mulch",
  range = readxl::cell_cols("A:E"),
  col_types = c(
    "date",
    "text",
    "text",
    "text",
    "numeric"
  )
)

if (ncol(raw) != 5L) {
  stop(
    "The worksheet must contain the expected first five columns.",
    call. = FALSE
  )
}

names(raw) <- c(
  "harvest_date",
  "block",
  "treatment",
  "row_position",
  "fresh_weight_g"
)

clean <- raw %>%
  mutate(
    harvest_date = as.Date(harvest_date),
    across(
      c(block, treatment, row_position),
      ~ na_if(toupper(trimws(.x)), "")
    )
  ) %>%
  filter(if_any(everything(), ~ !is.na(.x)))

# Check identifiers before filtering to avoid silent exclusions.
if (anyNA(clean[c("block", "treatment", "row_position")])) {
  stop(
    "Missing block, treatment, or row identifiers were found.",
    call. = FALSE
  )
}

if (any(!clean$treatment %in% c("T1", "T2", "T3", "T4"))) {
  stop(
    "Unexpected treatment identifiers were found.",
    call. = FALSE
  )
}

valid_rows <- c(
  "E", "EAST",
  "M", "MIDDLE", "CENTER", "CENTRE",
  "W", "WEST"
)

if (any(!clean$row_position %in% valid_rows)) {
  stop(
    "Unexpected row-position labels were found.",
    call. = FALSE
  )
}

# Use middle-row records for the three mowing treatments.
dat <- clean %>%
  filter(
    row_position %in% c("M", "MIDDLE", "CENTER", "CENTRE"),
    treatment %in% c("T2", "T3", "T4")
  )

if (nrow(dat) == 0L) {
  stop(
    "No middle-row records were found for T2, T3, and T4.",
    call. = FALSE
  )
}

if (anyNA(dat)) {
  stop(
    "Missing dates or weights were found in the selected records.",
    call. = FALSE
  )
}

if (any(!is.finite(dat$fresh_weight_g)) ||
    any(dat$fresh_weight_g <= 0)) {
  stop(
    "Selected fresh weights must be finite and greater than zero.",
    call. = FALSE
  )
}

if (any(!dat$block %in% c("B1", "B2", "B3", "B4"))) {
  stop(
    "Expected block identifiers B1, B2, B3, and B4.",
    call. = FALSE
  )
}

# ============================================================
# Define treatment order and labels
# ============================================================

# T4: Mowing only before broccoli transplanting.
# T3: Mowing once every two months.
# T2: Mowing once per month.

treatment_order <- c("T4", "T3", "T2")

treatment_labels <- c(
  T4 = "Pre-transplant\nonly (T4)",
  T3 = "Every 2 months\n(T3)",
  T2 = "Monthly\n(T2)"
)

treatment_colours <- c(
  T4 = "#D16B32",
  T3 = "#168C79",
  T2 = "#2877B7"
)

dat <- dat %>%
  mutate(
    block = factor(
      block,
      levels = c("B1", "B2", "B3", "B4")
    ),
    treatment = factor(
      treatment,
      levels = treatment_order
    )
  )

# Missing records must not automatically be treated as zero yield.
if (any(table(dat$block, dat$treatment) == 0L)) {
  stop(
    "At least one Block x Treatment combination has no records.",
    call. = FALSE
  )
}

last_harvest_date <- max(dat$harvest_date)

date_caption <- paste(
  "Data through",
  format(last_harvest_date, "%Y-%m-%d")
)

# ============================================================
# Summarize each experimental plot
# ============================================================

# Each Block x Treatment combination is one experimental plot.
# Pool all recorded harvest dates within each plot.
# All recorded heads are included, regardless of marketability.
# Areas are equal, but their exact sizes are unknown.
# Report fresh weight per sampled middle row, not per square metre.

plot_data <- dat %>%
  group_by(block, treatment) %>%
  summarise(
    total_kg = sum(fresh_weight_g) / 1000,
    head_count = n(),
    mean_head_g = mean(fresh_weight_g),
    .groups = "drop"
  )

metric_labels <- c(
  "A  Harvested fresh weight\n(kg / middle row)",
  "B  Harvested heads\n(number / middle row)",
  "C  Mean head fresh weight\n(g / harvested head)"
)

plot_long <- plot_data %>%
  pivot_longer(
    cols = c(total_kg, head_count, mean_head_g),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric = factor(
      metric,
      levels = c("total_kg", "head_count", "mean_head_g"),
      labels = metric_labels
    )
  )

# Calculate means and standard errors across plots.
# Each plot receives equal weight.

summary_data <- plot_long %>%
  group_by(treatment, metric) %>%
  summarise(
    n_plots = n(),
    mean_value = mean(value),
    se_value = sd(value) / sqrt(n()),
    .groups = "drop"
  )
```

## Harvest comparison

Only middle-row records are included. Each treatment has four experimental
plots. Coloured points represent individual plots. Black diamonds show
treatment means, and error bars show one standard error across plots.
These are descriptive summaries.

```{r harvest-comparison, fig.width=15, fig.height=6.5}
p_compare <- ggplot(
  plot_long,
  aes(x = treatment, y = value)
) +
  geom_point(
    aes(colour = treatment, shape = block),
    position = position_dodge(width = 0.55),
    size = 3.2,
    alpha = 0.9
  ) +
  geom_errorbar(
    data = summary_data,
    inherit.aes = FALSE,
    aes(
      x = treatment,
      ymin = mean_value - se_value,
      ymax = mean_value + se_value
    ),
    width = 0.14,
    linewidth = 0.7,
    colour = "black"
  ) +
  geom_point(
    data = summary_data,
    inherit.aes = FALSE,
    aes(x = treatment, y = mean_value),
    shape = 18,
    size = 4.2,
    colour = "black"
  ) +
  facet_wrap(
    ~ metric,
    scales = "free_y",
    nrow = 1
  ) +
  scale_x_discrete(labels = treatment_labels) +
  scale_colour_manual(
    values = treatment_colours,
    guide = "none"
  ) +
  scale_shape_manual(
    values = c(B1 = 16, B2 = 17, B3 = 15, B4 = 8)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0.02, 0.12))
  ) +
  expand_limits(y = 0) +
  labs(
    title = "Broccoli harvest across mowing regimes",
    subtitle = paste(
      date_caption,
      "| Middle rows only | 4 plots per treatment"
    ),
    x = "Cover crop mowing regime",
    y = NULL,
    shape = "Block",
    caption = paste0(
      "Coloured points = individual plots; ",
      "black diamonds = means; error bars = SE.\n",
      "All recorded heads are included. ",
      "Head-weight means give equal weight to each plot."
    )
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 20),
    plot.subtitle = element_text(
      size = 11,
      margin = margin(b = 18)
    ),
    strip.background = element_blank(),
    strip.text = element_text(
      face = "bold",
      size = 12,
      margin = margin(b = 12)
    ),
    axis.text.x = element_text(
      size = 10,
      margin = margin(t = 8)
    ),
    axis.title.x = element_text(
      margin = margin(t = 14)
    ),
    panel.spacing = grid::unit(1.2, "cm"),
    panel.grid.major.y = element_line(
      colour = "#E5E9EE",
      linewidth = 0.4
    ),
    legend.position = "bottom",
    plot.caption = element_text(hjust = 0, size = 9),
    plot.margin = margin(15, 20, 15, 15)
  )

print(p_compare)
```

## Cumulative harvest

Thick lines show treatment means across four plots. Thin lines show
individual plots. Dates without entries add no additional recorded weight;
this does not establish that no harvest occurred on those dates.

```{r cumulative-harvest, fig.width=11, fig.height=6}
daily_harvest <- dat %>%
  group_by(block, treatment, harvest_date) %>%
  summarise(
    daily_kg = sum(fresh_weight_g) / 1000,
    .groups = "drop"
  )

date_grid <- tidyr::expand_grid(
  block = levels(dat$block),
  treatment = treatment_order,
  harvest_date = seq(
    min(dat$harvest_date),
    last_harvest_date,
    by = "day"
  )
)

cumulative_data <- date_grid %>%
  left_join(
    daily_harvest,
    by = c("block", "treatment", "harvest_date")
  ) %>%
  mutate(daily_kg = coalesce(daily_kg, 0)) %>%
  arrange(block, treatment, harvest_date) %>%
  group_by(block, treatment) %>%
  mutate(cumulative_kg = cumsum(daily_kg)) %>%
  ungroup()

cumulative_means <- cumulative_data %>%
  group_by(treatment, harvest_date) %>%
  summarise(
    mean_kg = mean(cumulative_kg),
    .groups = "drop"
  )

p_cumulative <- ggplot() +
  geom_step(
    data = cumulative_data,
    aes(
      x = harvest_date,
      y = cumulative_kg,
      colour = treatment,
      group = interaction(block, treatment)
    ),
    linewidth = 0.5,
    alpha = 0.25
  ) +
  geom_step(
    data = cumulative_means,
    aes(
      x = harvest_date,
      y = mean_kg,
      colour = treatment,
      group = treatment
    ),
    linewidth = 1.2
  ) +
  scale_colour_manual(
    values = treatment_colours,
    breaks = treatment_order,
    labels = unname(treatment_labels[treatment_order])
  ) +
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%b %d"
  ) +
  scale_y_continuous(
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.06))
  ) +
  labs(
    title = "Cumulative broccoli harvest",
    subtitle = date_caption,
    x = "Harvest date",
    y = "Cumulative fresh weight (kg / middle row)",
    colour = "Mowing regime",
    caption = paste(
      "Middle rows only.",
      "Thick lines = treatment means;",
      "thin lines = individual plots."
    )
  ) +
  theme_classic(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    legend.position = "bottom",
    panel.grid.major.y = element_line(
      colour = "#E5E9EE",
      linewidth = 0.4
    ),
    plot.caption = element_text(hjust = 0, size = 9)
  )

print(p_cumulative)
```

## Treatment summary

Means and standard errors are calculated across four plots per treatment.

```{r treatment-summary}
display_summary <- summary_data %>%
  transmute(
    Treatment = as.character(treatment),
    Measurement = gsub("\n", " ", as.character(metric)),
    Mean = mean_value,
    SE = se_value,
    Plots = n_plots
  )

knitr::kable(
  display_summary,
  digits = 2,
  row.names = FALSE
)
```

## Plot summary

Each row represents one experimental plot.

```{r plot-summary}
display_plots <- plot_data %>%
  transmute(
    Block = block,
    Treatment = treatment,
    `Fresh weight (kg)` = total_kg,
    `Harvested heads` = head_count,
    `Mean head weight (g)` = mean_head_g
  )

knitr::kable(
  display_plots,
  digits = 2,
  row.names = FALSE
)
```

```{r save-results, include=FALSE}
# Save both figures as PNG and PDF.

figures <- list(
  Broccoli_mowing_comparison = p_compare,
  Broccoli_cumulative_harvest = p_cumulative
)

for (figure_name in names(figures)) {
  is_comparison <- figure_name == "Broccoli_mowing_comparison"

  figure_width <- if (is_comparison) 15 else 11
  figure_height <- if (is_comparison) 6.5 else 6

  for (extension in c("png", "pdf")) {
    ggsave(
      filename = file.path(
        output_dir,
        paste0(figure_name, ".", extension)
      ),
      plot = figures[[figure_name]],
      width = figure_width,
      height = figure_height,
      units = "in",
      dpi = 300,
      bg = "white"
    )
  }
}

# Save numerical summaries.

write.csv(
  plot_data,
  file.path(output_dir, "Plot_summary.csv"),
  row.names = FALSE
)

write.csv(
  summary_data,
  file.path(output_dir, "Treatment_summary.csv"),
  row.names = FALSE
)

write.csv(
  cumulative_data,
  file.path(output_dir, "Cumulative_harvest_by_plot.csv"),
  row.names = FALSE
)
```